### 2020-08-21 latex/vlp.tex
Better motivaated why VNHCss are important and why we study the VLP

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical dynamical system, which is often
used to model a the motion of a person on a standing swing
\cite{dynamics_periodic_vlp,pumping_swing_standing_squatting,how_to_pump_a_swing}.
% TODO: Why else is it useful? Apparently it models cranes, satellites, etc
% TODO: Motivate more why this is a good problem to study

The variable length pendulum (VLP) is modelled as a point with mass \(m\)
connected to a pivot by a massless rod of varying length \(l\) and angle 
\(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
\ref{fig:vlp_model}. 
In a realistic VLP, the rod length \(l\) varies between some minimum
length \(\underline{l} \geq 0\) and some maximum length 
\(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).

\begin{figure}
   \centering
   \includestandalone[width=0.4\textwidth]{images/vlp_model}
   \caption{The variable length pendulum with a mass at the
   tip.}\label{fig:vlp_model}
\end{figure}

@@

What this means is we can treat \(l\) as the control input directly, rather than
modelling it as a configuration variable. Re-deriving the Hamiltonian and the
dynamics with this in mind, we get the system 
(\ref{eqn:vlp-hamiltonian}) with phase \((q,p) \in \mathbb{S}^1 \times \R\). 
Note that the control input \(l(t)\) and its derivative \(\dot{l}(t)\) are both
known variables.
\begin{align}\label{eqn:vlp-hamiltonian}
   \mathcal{H}(q,p) &= \frac{p^2}{ml^2} - \frac{1}{2}\dot{l}^2 - mgl\cos(q) \\
     &\begin{cases}
        \dot{q} = \frac{p}{ml^2} \\
        \dot{p} = -mgl\sin(q) \\
      \end{cases}\nonumber
\end{align}
In particular, the Hamiltonian of this simplified model is no longer equal to
the total mechanical energy of the system, which is given by
(\ref{eqn:vlp-energy}).
\begin{equation}\label{eqn:vlp-energy}
   E(q,p) = \frac{p^2}{ml^2} + \frac{1}{2}\dot{l}^2 - mgl\cos(q)
\end{equation}
\section{The VLP Constraint}
% TODO: Go through the development of the constraint and prove it gains energy
% TODO: Explain how we came up with the constraint first. Talk about the child
% on a swing, the paper showing the time-optimal control input, and describe why
% we want to remove time from the equation. Discuss the lmin and lmax choices.
% Why do we need to use a VNHC instead of a VHC: we are modelling the
% time-optimal controller in a time-independent way, which requires knowledge of
% which quadrant in (q,p)-space we're in. This means using the angle theta,
% which is inherently a VNHC.

>>>>

%/========== Variable Length Pendulum ==========/%
\chapter{Application of VNHCS: The Variable Length Pendulum}\label{sec:vlp}
\section{Motivation}
The variable length pendulum (VLP) is a classical underactuated dynamical system
which is often used to model the motion of a person on a swing
\cite{pumping_swing_standing_squatting,how_to_pump_a_swing}.
The VLP also represents the swinging of a crane, the (simplified) motion of a
gymnast on a bar \cite{pendulum_length_giant_gymnastics}, and the
tuned-mass-damper systems which stabilize skyscrapers
\cite{vlp_tuned_mass_damper}, among others.

The motion of the VLP has been well studied (see for instance
\cite{dynamics_periodic_vlp}, and many control mechanisms exist
to stabilize trajectories of the system. While many of these control mechanism
are time-dependent, \cite{vlp_energy_shaping}
offers a time-independent technique to inject energy into the system and
stabilize desired energy level sets. The authors designed a controller through a
technique called \textit{energy shaping}, and proved that their control
mechanism would allow the VLP to achieve any desired energy level set.
However, the energy injection mechanism is
ad-hoc in the sense that it is not derived from any human-like behaviour. In
this chapter, we will use VNHCs to provably stabilize energy level sets as well;
the difference is that the control mechanism we derive will stay true to human
motion and can be implemented by any person on a swing.

\section{Dynamics of the Variable Length Pendulum}
% Derive the dynamics of the VLP in Hamiltonian form
We will model the VLP as a point mass \(m\)
connected to a fixed pivot by a massless rod of varying length \(l\) with angle 
\(q \in \mathbb{S}^1\) from the vertical, as is seen in Figure
\ref{fig:vlp_model}. 
We will also ignore any damping and frictional forces in this model.
In a realistic VLP, the rod length \(l\) varies between some minimum
length \(\underline{l} \geq 0\) and some maximum length 
\(\overline{l} > \underline{l}\). The configuration of the VLP is the vector
\(\mathbf{q} := (q,l) \in \mathbb{S}^1 \times [\underline{l},\overline{l}]\).

\begin{figure}
   \centering
   \includestandalone[width=0.4\textwidth]{images/vlp_model}
   \caption{The representation of the variable length pendulum as a mass at the
      tip of a massless rod.}\label{fig:vlp_model}
\end{figure}

@@

What this means is we can treat \(l\) as the control input directly, rather than
modelling it as a configuration variable. Re-deriving the Hamiltonian and the
dynamics with this in mind, we get the system 
(\ref{eqn:vlp-hamiltonian}) with phase \((q,p) \in \mathbb{S}^1 \times \R\). 
Note that the control input \(l(t)\) and its derivative \(\dot{l}(t)\) are both
known variables.
\begin{align}\label{eqn:vlp-hamiltonian}
   \mathcal{H}(q,p) &= \frac{p^2}{ml^2} - \frac{1}{2}\dot{l}^2 - mgl\cos(q) \\
     &\begin{cases}
        \dot{q} = \frac{p}{ml^2} \\
        \dot{p} = -mgl\sin(q) \\
      \end{cases}\nonumber
\end{align}
In particular, the Hamiltonian of this simplified model is no longer equal to
the total mechanical energy of the system, which is given by
(\ref{eqn:vlp-energy}).
\begin{equation}\label{eqn:vlp-energy}
   E(q,p) = \frac{p^2}{ml^2} + \frac{1}{2}\dot{l}^2 - mgl\cos(q)
\end{equation}

\section{The VLP Constraint}
% TODO: Go through the development of the constraint and prove it gains energy
% TODO: Explain how we came up with the constraint first. Talk about the child
% on a swing, the paper showing the time-optimal control input, and describe why
% we want to remove time from the equation. Discuss the lmin and lmax choices.
% Why do we need to use a VNHC instead of a VHC: we are modelling the
% time-optimal controller in a time-independent way, which requires knowledge of
% which quadrant in (q,p)-space we're in. This means using the angle theta,
% which is inherently a VNHC.

### 2020-08-21 latex/vnhcs.tex
Better motivaated why VNHCss are important and why we study the VLP

Many authors have attempted to extend these results to enforce a relation 
\(h(q,\dot{q}) = 0\) of the full state. Since these relations use actuators to
restrict both the configuration and generalized velocity, they are called
virtual \textit{nonholonomic} constraints. This idea has been used for
human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and improving bipedal locomotion \cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,nhvc_incline_walking,output_nhvc_bipedal_control}.

>>>>

Many authors have attempted to extend these results to enforce a relation 
\(h(q,\dot{q}) = 0\) of the full state. Since these relations use actuators to
restrict both the configuration and generalized velocity, they are called
virtual \textit{nonholonomic} constraints. This idea has been used for
human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and improving bipedal locomotion \cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Most interestingly, this nonholonomic approach has been shown to be more robust
than standard VHCs when applied to bipedal robotics \cite{nhvc_incline_walking}.
This suggests that virtual nonholonomic constraints are more capable of
stabilizing specific energy levels while rejecting disturbances, all while
producing realistic biological motion. It is for this reason we choose to study
nonholonomic constraints in this thesis.
